---
title: "New York City Crime & Population Trends"
author: "Brent Brewington (github.com/bbrewington)"
date: "October 2, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)      # to scrape wikipedia data
library(readr)      # to read csv data
library(ggplot2)    # plotting (I'm an addict)
library(dplyr)      # data munging
library(tidyr)      # data tidying
library(scales)     # package for tidying up plot axes
```


#### Read crime & population data into R:


##### Scrape NYC Population Data from Wikipedia
```{r scrape_population_data, echo = TRUE, message = FALSE}
nyc.population <- read_html("https://en.wikipedia.org/wiki/Demographics_of_New_York_City") %>% 
     html_nodes(".toccolours") %>% html_table() %>% as.data.frame() %>% 
     .[2:32,1:2] %>% 
     select(year = Historical.population, population = Historical.population.1) %>%     
     mutate(year = as.numeric(year), population = parse_number(population))
glimpse(nyc.population)
```


##### Get Crime Data from nyc.gov/html/nypd:

1. Downloaded Excel (.xls) file: http://www.nyc.gov/html/nypd/downloads/excel/analysis_and_planning/seven_major_felony_offenses_2000_2015.xls
2. Copy+Paste data into csv file
3. Read into R via readr::read_csv
```{r read_crime_data, echo = TRUE, message = FALSE}
nyc.crime <- read_csv("nyc_crimedata_2000-2015.csv", col_names = TRUE) %>% 
     gather(year, crime.count, `2000`:`2015`) %>% 
     mutate(year = as.numeric(year))
glimpse(nyc.crime)
```

##### Plot of NYC population trend
```{r nyc_population_trend, echo = FALSE}
ggplot(nyc.population, aes(year, population)) +
     geom_point() + geom_line() + 
     scale_y_continuous(labels = comma_format()) + 
     ggtitle("NYC Population Over Time") + ylab("Population") + xlab("Year")
```

##### Fit linear model to NYC population (1980-2010)
```{r nyc_population_linear_model, echo = FALSE}
nyc.population1 <- nyc.population %>% filter(year >= 1980)
nyc.population_model1 <- lm(nyc.population1$population ~ nyc.population1$year)
summary(nyc.population_model1)
nyc.population1 %>% mutate(population.est = -75428524 + 41646 * year) %>%
     ggplot(aes(x = year)) + geom_line(aes(y = population.est), color = "red") + geom_point(aes(y = population)) + ggtitle("Fitted Line: NYC Population from 1980-2010 (by decade)") +
     xlab("Year") + ylab("Population") + scale_y_continuous(labels = comma_format())
```

NYC population from 1980-2010 = -75428524 + (41646 * year)

##### Create dataset: NYC Crime Rate 2000-2015
```{r nyc_crime_rate_dataset}
nyc.population_crime <- data.frame(year = 2000:2015) %>% 
     mutate(population.est = -75428524 + 41646 * year) %>%
     left_join(nyc.crime, by = "year") %>%
     mutate(crime.rate = crime.count / population.est) %>% 
     mutate(crime.rate_per.million = crime.rate * 1000000)
```

```{r display_nyc_crime_rate_dataset, echo = FALSE}
glimpse(nyc.population_crime)
```

##### Create dataset: NYC Violent Crime Rate 2000-2015
```{r nyc_population_violentcrime_dataset}
nyc.population_violentcrime <- data.frame(year = 2000:2015) %>% 
     mutate(population.est = -75428524 + 41646 * year) %>%
     left_join(nyc.crime %>% filter(OFFENSE %in% c("MURDER & NON-NEGL. MANSLAUGHTER", "RAPE", "ROBBERY", "FELONY ASSAULT")) %>% group_by(year) %>% 
                    summarise(crime.count = sum(crime.count)), by = "year") %>%
     mutate(crime.rate = crime.count / population.est, 
            crime.rate_per.million = 1000000 * crime.count / population.est)
```

```{r display_nyc_population_violentcrime_dataset, echo = FALSE}
glimpse(nyc.population_violentcrime)
```

### Plots
```{r plot_seven_major_felonies}
ggplot(nyc.population_crime, aes(year, crime.rate_per.million, color = OFFENSE)) + geom_point() + geom_smooth(se = FALSE) + ggtitle("NYC Seven Major Felony Offenses by Year: 2000-2015") +
     xlab("Year") + ylab("Seven Major Felonies Per Million NYC Residents")
```

```{r plot_murder_rate}
ggplot(nyc.population_crime %>% filter(OFFENSE == "MURDER & NON-NEGL. MANSLAUGHTER"), aes(year, crime.rate_per.million)) + geom_point() + geom_smooth() + geom_line() +
     xlab("Year") + ylab("Murders Per Million NYC Residents") + ggtitle("NYC Murder Rate: 2000-2015") + geom_label(aes(x = 2010, y = 90), label = "Blue Line: Lowess Smoothing (with 95% confidence band)") + scale_y_continuous(limits = c(0, 100))
```

```{r plot_violent_crime}
ggplot(nyc.population_violentcrime, aes(year, crime.rate_per.million)) + geom_bar(stat = "identity", aes(fill = "Total Violent Crime Rate")) + geom_smooth(data = nyc.population_crime %>% filter(OFFENSE %in% c("MURDER & NON-NEGL. MANSLAUGHTER", "RAPE", "ROBBERY", "FELONY ASSAULT")), aes(x = year, y = crime.rate_per.million, color = OFFENSE), se = FALSE) + geom_point(data = nyc.population_crime %>% filter(OFFENSE %in% c("MURDER & NON-NEGL. MANSLAUGHTER", "RAPE", "ROBBERY", "FELONY ASSAULT")), aes(x = year, y = crime.rate_per.million, color = OFFENSE)) + scale_fill_manual(values = "black", name = NULL) + scale_color_discrete(name = NULL) +
     ggtitle("NYC Violent Crime Rate: 2000-2015") + ylab("Violent Crimes per Million NYC Residents") + xlab("Year")
```